\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-10-29 doc/flexiblesusy-paper.tex
fix typo theses --> these.
added more details on testing with SUSY-HITS.

The standard input and output of the generated code is given in SLHA
format and an extension of this to incorporate new blocks for models
which go beyond the MSSM and NMSSM. We use the well known SLHAea
\cite{slhaea} to create these SLHA files and the extra blocks may be chosen
by the user. Since the SLHA standard is intended precisely for the
purpose, MSSM and NMSSM codes should be compliant with this, and we
have tested that HDECAY\cite{hdecay} can understand the SLHA output of
FlexibleSUSY.

@@

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by theses equations.  For
example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
\fs handles only real model parameters and renormalizable theories,
which restricts the number of possible multiple solutions of Eqs.\
\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
number of equations.  In some cases, as for example in the real MSSM,
% where $\mu$ is chosen to be fixed by the EWSB equations
these different solutions might be related by one or more free global
signs in the form
%
\begin{align}
  \mu = \sign\mu \cdot \sqrt{\ldots} .
\end{align}
%
\fs recognizes such cases and introduces these global signs as
additional free parameters.  The case of the real MSSM, where only
$|\mu|$ is fixed by the EWSB equations, is therefore handled
automatically.  If, however, the different solutions are not related
by global signs, then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
is chosen to be fixed by the EWSB equations: There the tree-level
solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  In this model the user can parametrize
the tree-level solution in terms of $\sign X$ in the
\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
list of free input parameters, see
\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
user decides to not pick a particular solution by leaving the variable
\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
the tree-level EWSB equations numerically via an iteration.  In this
case, however, the user does not have the option to chose between the
different solutions.

@@

\bibitem{hdecay}  A. Djouadi, M. Spira and P.M. Zerwas, Phys. Lett.  B {\bf 264}
(1991) 440 and  Z. Phys. C {\bf 70} (1996) 427;  M. Spira {\it et al.}, Nucl.\
Phys.\ B {\bf 453} (1995) 17;  A.~Djouadi, J.~Kalinowski and M.~Spira, Comput.
Phys. Commun. {\bf 108} (1998) 56; J.~M.~Butterworth, A.~Arbey, L.~Basso, S.~Belov, A.~Bharucha, F.~Braam, A.~Buckley and M.~Campanelli {\it et al.},
  %``The Tools and Monte Carlo working group Summary Report,''
  arXiv:1003.1643 [hep-ph].

>>>>

The standard input and output of the generated code is given in
SLHA-style format where the user has control of the Block names by
editing the SARAH files and may also add extra blocks in the
FlexibleSUSY.m model file which is decribed in section 5. We use the
well known SLHAea \cite{slhaea} to create these SLHA files and with a
simple modifications these can be passed directly to another code
which uses them as an alternative to writing out the SLHA file. Since the
SLHA standard is intended precisely for the purpose, MSSM and NMSSM
codes should be compliant with this.  We have tested that the versions
of HDECAY\cite{hdecay} and SDECAY\cite{sdecay} inside SUSY-HIT
\cite{susyhit} can understand the SLHA output of FlexibleSUSY and give
matching output to that from SOFTSUSY.

@@

\fs assumes that each SUSY model contains Higgs bosons, which trigger
a spontaneous breaking of the electroweak symmetry.  The corresponding
EWSB consistency conditions are formulated in \fs at the one-loop
level as
%
\begin{align}
  0 = \frac{\partial V^\text{tree}}{\partial v_i} - t_i,
  \label{eq:one-loop-ewsb-eq}
\end{align}
%
where $V^\text{tree}$ is the tree-level Higgs potential, $v_i$ is the
VEV corresponding to the Higgs fields $H_i$ and $t_i$ is the one-loop
tadpole diagram of $H_i$.  Already at the tree-level ($t_i = 0$) Eqs.\
\eqref{eq:one-loop-ewsb-eq} can have multiple solutions, depending on
which paramters are chosen to be fixed by these equations.  For
example in the real MSSM Eqs.\ \eqref{eq:one-loop-ewsb-eq} can only
determine $|\mu|$, while the sign of $\mu$ is not fixed.  Currently
\fs handles only real model parameters and renormalizable theories,
which restricts the number of possible multiple solutions of Eqs.\
\eqref{eq:one-loop-ewsb-eq} to be less or equal to four times the
number of equations.  In some cases, as for example in the real MSSM,
% where $\mu$ is chosen to be fixed by the EWSB equations
these different solutions might be related by one or more free global
signs in the form
%
\begin{align}
  \mu = \sign\mu \cdot \sqrt{\ldots} .
\end{align}
%
\fs recognizes such cases and introduces these global signs as
additional free parameters.  The case of the real MSSM, where only
$|\mu|$ is fixed by the EWSB equations, is therefore handled
automatically.  If, however, the different solutions are not related
by global signs, then \fs writes all solutions to the file
\code{models/<model>/<model>_tree_level_EWSB_solution.m}.  The user
has then the option to pick a particular solution by setting it in the
\code{TreeLevelEWSBSolution} variable in the model file.  This is for
example done in the $Z_3$-violating NMSSM (\code{SMSSM}), where $\mu$
is chosen to be fixed by the EWSB equations: There the tree-level
solution for $\mu$ has the form
%
\begin{align}
  \mu = -\frac{v_s \lambda}{\sqrt{2}} + \sign X \cdot \sqrt{\ldots}
\end{align}
%
with a non-global $\sign X$.  In this model the user can parametrize
the tree-level solution in terms of $\sign X$ in the
\code{TreeLevelEWSBSolution} variable and add $\sign X$ by hand to the
list of free input parameters, see
\code{model_files/SMSSM/FlexibleSUSY.m.in} for an example.  If the
user decides to not pick a particular solution by leaving the variable
\code{TreeLevelEWSBSolution} empty, \fs tries to find a solution to
the tree-level EWSB equations numerically via an iteration.  In this
case, however, the user does not have the option to chose between the
different solutions.

@@

\bibitem{hdecay}  A. Djouadi, M. Spira and P.M. Zerwas, Phys. Lett.  B {\bf 264}
(1991) 440 and  Z. Phys. C {\bf 70} (1996) 427;  M. Spira {\it et al.}, Nucl.\
Phys.\ B {\bf 453} (1995) 17;  A.~Djouadi, J.~Kalinowski and M.~Spira, Comput.
Phys. Commun. {\bf 108} (1998) 56; J.~M.~Butterworth, A.~Arbey, L.~Basso, S.~Belov, A.~Bharucha, F.~Braam, A.~Buckley and M.~Campanelli {\it et al.},
  %``The Tools and Monte Carlo working group Summary Report,''
  arXiv:1003.1643 [hep-ph].
%\cite{Muhlleitner:2003vg}
\bibitem{sdecay} 
  M.~Muhlleitner, A.~Djouadi and Y.~Mambrini,
  %``SDECAY: A Fortran code for the decays of the supersymmetric particles in the MSSM,''
  Comput.\ Phys.\ Commun.\  {\bf 168}, 46 (2005)
  [hep-ph/0311167].
  %%CITATION = HEP-PH/0311167;%%
  %235 citations counted in INSPIRE as of 29 Oct 2014

%\cite{Djouadi:2006bz}
\bibitem{susyhit} 
  A.~Djouadi, M.~M.~Muhlleitner and M.~Spira,
  %``Decays of supersymmetric particles: The Program SUSY-HIT (SUspect-SdecaY-Hdecay-InTerface),''
  Acta Phys.\ Polon.\ B {\bf 38}, 635 (2007)
  [hep-ph/0609292].
  %%CITATION = HEP-PH/0609292;%%
  %238 citations counted in INSPIRE as of 29 Oct 2014
